{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-07-13 sftrends_manu.Rnw
d7 and c10 trend tables added

% trends in sio and chl at d7
<<chlsio>>=

data(diat_dat)

sioraw <- filter(diat_dat, resvar %in% 'sio2') %>% 
  .$mod %>% 
  .[[1]]
chlraw <- filter(diat_dat, resvar %in% 'chl') %>% 
  .$mod %>% 
  .[[1]]

@@

# summarize sio2 and chl trends
sio <- wrtdstrnd(sioraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)

# combine for trends for table
totab <- rbind(sio, chl) %>% 
  mutate(
    chg = round(chg, 1), 
    chg = ifelse(chg < 0, chg, paste0('\\textit{\\textbf{', chg , '}}')), 
    ave = as.character(round(ave, 1))
    ) %>% 
  gather('var', 'val', chg:ave) %>% 
  unite('col', res, var) %>% 
  spread(col, val) %>% 
  mutate(cat = factor(cat, levels = c('1976-2013', yrlabs, molabs), labels = c('1976-2013', yrlabs, molabs))) %>% 
  arrange(cat) %>% 
  data.frame(sep = c(rep('Annual', 5), rep('Seasonal', 4)), .)

# final table formatting
segs <- c('All', 'Annual', 'Seasonal')
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('Chl-\\textit{a}', 'SiO$_2$')
names(totab) <- rep(c('Mean', '\\% change'), 2)

cap.val<-'Summaries of flow-normalized trends in chlorophyll ($\\mu$g $L^{-1}$) and silicon dioxide (mg $L^{-1}$) concentrations for different time periods. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(1, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2),
  rowname = cats,
  label = 'tab:chlsio'
  )

>>>>

% trends in din, chl, sio2 at d7
<<d7chg>>=

data(h3dat)

d7dat <- filter(h3dat, Site_Code %in% 'D7')

dinraw <- filter(d7dat, resvar %in% 'din') %>% 
  .$mod %>% 
  .[[1]]
chlraw <- filter(d7dat, resvar %in% 'chl') %>% 
  .$mod %>% 
  .[[1]]
sioraw <- filter(d7dat, resvar %in% 'sio2') %>% 
  .$mod %>% 
  .[[1]]

@@

# summarize din, sio2, chl trends
din <- wrtdstrnd(dinraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
din <- wrtdstrnd(dinraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., din) %>% 
  data.frame(res = 'din', .)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)

# combine for trends for table
totab <- rbind(din, chl, sio) %>% 
  mutate(
    chg = round(chg, 1), 
    chg = ifelse(chg < 0, chg, paste0('\\textit{\\textbf{', chg , '}}')), 
    ave = as.character(round(ave, 1))
    ) %>% 
  gather('var', 'val', chg:ave) %>% 
  unite('col', res, var) %>% 
  spread(col, val) %>% 
  mutate(cat = factor(cat, levels = c('1976-2013', yrlabs, molabs), labels = c('1976-2013', yrlabs, molabs))) %>% 
  arrange(cat) %>% 
  data.frame(sep = c(rep('Annual', 5), rep('Seasonal', 4)), .) %>% 
  select(sep, cat, din_ave, din_chg, chl_ave, chl_chg, sio_ave, sio_chg)

# final table formatting
segs <- c('All', 'Annual', 'Seasonal')
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('DIN', 'Chl-\\textit{a}', 'SiO$_2$')
names(totab) <- rep(c('Mean', '\\% change'), 3)

cap.val<-'Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(1, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2, 2),
  rowname = cats,
  label = 'tab:d7chg'
  )

% trends in din, chl, sio2 at c10
<<c10chg>>=

data(h3dat)

c10dat <- filter(h3dat, Site_Code %in% 'C10')

dinraw <- filter(c10dat, resvar %in% 'din') %>% 
  .$mod %>% 
  .[[1]]
chlraw <- filter(c10dat, resvar %in% 'chl') %>% 
  .$mod %>% 
  .[[1]]
sioraw <- filter(c10dat, resvar %in% 'sio2') %>% 
  .$mod %>% 
  .[[1]]

# summarize din, sio2, chl trends
din <- wrtdstrnd(dinraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
din <- wrtdstrnd(dinraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., din) %>% 
  data.frame(res = 'din', .)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)

# combine for trends for table
totab <- rbind(din, chl, sio) %>% 
  mutate(
    chg = round(chg, 1), 
    chg = ifelse(chg < 0, chg, paste0('\\textit{\\textbf{', chg , '}}')), 
    ave = as.character(round(ave, 1))
    ) %>% 
  gather('var', 'val', chg:ave) %>% 
  unite('col', res, var) %>% 
  spread(col, val) %>% 
  mutate(cat = factor(cat, levels = c('1976-2013', yrlabs, molabs), labels = c('1976-2013', yrlabs, molabs))) %>% 
  arrange(cat) %>% 
  data.frame(sep = c(rep('Annual', 5), rep('Seasonal', 4)), .) %>% 
  select(sep, cat, din_ave, din_chg, chl_ave, chl_chg, sio_ave, sio_chg)

# final table formatting
segs <- c('All', 'Annual', 'Seasonal')
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('DIN', 'Chl-\\textit{a}', 'SiO$_2$')
names(totab) <- rep(c('Mean', '\\% change'), 3)

cap.val<-'Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station C10. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(1, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2, 2),
  rowname = cats,
  label = 'tab:c10chg'
  )

### 2016-07-13 sftrends_manu.tex
d7 and c10 trend tables added

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Thu Jun 23 14:00:35 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

% trends in sio and chl at d7
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2), rowname = cats,     label = "tab:chlsio")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in chlorophyll ($\mu$g $L^{-1}$) and silicon dioxide (mg $L^{-1}$) concentrations for different time periods. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:chlsio}} 
\begin{center}
\begin{tabular}{lllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&\tabularnewline
~~1976-2013&3.5&-80.1&&12.5&\textit{\textbf{33.6}}\tabularnewline
\hline
{\bfseries Annual}&&&&&\tabularnewline
~~1976-1985&7.8&-62.3&&10.5&\textit{\textbf{8.9}}\tabularnewline
~~1986-1994&2.2&-62.1&&12.4&\textit{\textbf{15.6}}\tabularnewline
~~1995-2003&1.6&\textit{\textbf{20.2}}&&13.7&\textit{\textbf{4.5}}\tabularnewline
~~2004-2013&2&\textit{\textbf{31.3}}&&13.6&-2\tabularnewline
\hline
{\bfseries Seasonal}&&&&&\tabularnewline
~~JFM&1.9&-43.2&&14.8&\textit{\textbf{7.3}}\tabularnewline
~~AMJ&4.3&-72.8&&12.9&\textit{\textbf{25.7}}\tabularnewline
~~JAS&4.8&-88.4&&11.1&\textit{\textbf{70.8}}\tabularnewline
~~OND&2.1&-79.4&&11.7&\textit{\textbf{41.7}}\tabularnewline
\hline
\end{tabular}\end{center}

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Wed Jul 13 15:18:11 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

% trends in din, chl, sio2 at d7
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:d7chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:d7chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&0.4&-4.2&&4.6&-82.4&&12.5&\textit{\textbf{30}}\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&0.4&-15.2&&10.2&-67.9&&10.7&\textit{\textbf{8.4}}\tabularnewline
~~1986-1994&0.4&\textit{\textbf{15.4}}&&3.3&-48.8&&12.3&\textit{\textbf{13.5}}\tabularnewline
~~1995-2003&0.5&\textit{\textbf{4.3}}&&2.1&-3.9&&13.5&\textit{\textbf{3.7}}\tabularnewline
~~2004-2013&0.5&-10.6&&2.5&\textit{\textbf{31.7}}&&13.7&-1.8\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~JFM&0.5&-4.9&&2.7&-56&&15&\textit{\textbf{4.4}}\tabularnewline
~~AMJ&0.4&\textit{\textbf{17.4}}&&6.1&-74.3&&12.9&\textit{\textbf{20.9}}\tabularnewline
~~JAS&0.3&\textit{\textbf{40.6}}&&6.2&-89&&10.9&\textit{\textbf{79.1}}\tabularnewline
~~OND&0.5&\textit{\textbf{13.1}}&&2.7&-83.8&&11.7&\textit{\textbf{33.7}}\tabularnewline
\hline
\end{tabular}\end{center}

% trends in din, chl, sio2 at c10
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:c10chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station C10. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:c10chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&1.3&-16.9&&27.6&-4.9&&14.5&-23.1\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&1.3&\textit{\textbf{16.6}}&&26.6&-20.5&&15.7&-5\tabularnewline
~~1986-1994&1.4&\textit{\textbf{6.6}}&&27.5&\textit{\textbf{23.7}}&&14.8&-3.9\tabularnewline
~~1995-2003&1.4&-1.1&&29.3&-15.2&&14.1&-5.1\tabularnewline
~~2004-2013&1.2&-32.6&&27.3&\textit{\textbf{11}}&&13&-9.6\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~JFM&1.4&-20.5&&12&\textit{\textbf{14.8}}&&14.3&-16.9\tabularnewline
~~AMJ&1&-26.4&&43&\textit{\textbf{3.1}}&&13.1&-27.5\tabularnewline
~~JAS&1.3&-3.7&&46.6&-13.6&&14.8&-26.1\tabularnewline
~~OND&1.6&-6.8&&8.7&-18.5&&15.6&-16\tabularnewline
\hline
\end{tabular}\end{center}
